SubsidyExplorer - Step 1: Data Wrangling

### 2020-08-06 scripts/01_data_wrangling.Rmd
finalizing data sources

```{r}
# Paths to subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-2019-fse", "FISH_FSE_10032020183349581.csv")

# Paths to demographic data files
fao_fishers_path <- here::here("data", "raw", "FAO-2019-yearbook", "fao_fishers_yearbook_070619.csv")
teh_fte_path <- here::here("data", "raw", "Teh-and-Sumaila-2013-fte", "Fisher employment FTE.xlsx")

# Paths to fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-2019-capture-production", "Capture_2019.1.0/")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")

# Paths to fisheries mgmt files 
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
melnychuk_fmi_path <- here::here("data", "raw", "Melnychuk-et-al-2017-fmi", "fmi_values_country.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-2018-stock-assessment-database", "RLSADB v4.40", "DB Files With Assessment Data", "RLSADB v4.40 (assessment data only).xlsx")
anderson_fpi_path <- here::here("data", "raw", "Anderson-et-al-2015-fpi", "fpi_values_fishery.csv")

# Paths to other files
iuu_vessel_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")
cap_tier_path <- here::here("data", "raw", "USA-Cap-Tier-Proposal", "US_cap_tier_data.csv")
```

@@

**Fisheries (Subsidy) Data**
Estimates of annual fisheries subsidies provision by state from Sumaila et al. (2019) in 2018 US\$. These data were provided to us directly by the authors as a .xlsx file and are stored at `r sumaila_subsidies_path`.
> Sumaila, U. R., Ebrahim, N., Schuhbauer, A., Skerritt, D., Li, Y., Kim, H. S., … Pauly, D. (2019). Updated estimates and analysis of global fisheries subsidies. Marine Policy, 109, 103695. https://doi.org/10.1016/j.marpol.2019.103695

Estimates of fisheries support provision by state from the OECD (2008 - 2018). These data were last downloaded from the OECD iLibrary portal as a CSV file on March 10, 2019 and are stored as `r oecd_fse_path`.
> OECD. (2019). Fisheries: Fisheries support estimates. Fisheries and Aquaculture Statistics, OECD Agriculture Statistics (Database). https://doi.org/10.1787/ade64fdc-en

@@

Total number of fishers by state from the FAO Yearbook of Fishery and Aquaculture Statistics (misc. years between 1995 and 2017). These data were manually transcribed from PDF (`./data/raw/fleet_&_employment.pdf`) to CSV format on July 6, 2019 and are stored as `r fao_fishers_path`. 
> FAO (2019). FAO yearbook. Fishery and Aquaculture Statistics 2017, http://www.fao.org/fishery/statistics/global-fishers/en (accessed on 24 October 2019).

Estimated number of full-time jobs in fisheries from Teh and Sumaila (2013) in 2003. These data were provided to us directly by the authors as a .xlsx file and are stored as `r teh_fte_path`. 
> Teh, L. C. L., & Sumaila, U. R. (2011). Contribution of marine fisheries to worldwide employment. Fish and Fisheries, 14(1), 77–88. https://doi.org/10.1111/j.1467-2979.2011.00450.x

**Other Fisheries Data**
Fisheries landings in 2017 from the FAO Global Capture Production database (2019) and estimates of landed value generated using the ex-vessel price database created by Melnychuk et al. (2017). These data were were wrangled in the previous script, and the edited data are stored as `r catch_dat_path`. 
> FAO (2019). FAO yearbook. Fishery and Aquaculture Statistics 2017, http://www.fao.org/fishery/statistics/global-capture-production/en (accessed on 24 October 2019).

Estimated landed price 
> Melnychuk, M. C., Clavelle, T., Owashi, B., & Strauss, K. (2017). Reconstruction of global ex-vessel prices of fished species. ICES Journal of Marine Science, 74(1), 121–133. https://doi.org/10.1093/icesjms/fsw169

RAM Stock assessment database
> RAM Legacy Stock Assessment Database. (2018). Version 4.44-assessment-only. Released 2018-12-22. DOI:10.5281/zenodo.2542919

Stock status database
> Costello, C., Ovando, D., Clavelle, T., Strauss, C. K., Hilborn, R., Melnychuk, M. C., Branch, T. A., Gaines, S. D., Szuwalski, C. S., Cabral, R. B., Rader, D. N., & Leland, A. (2016). Global fishery prospects under contrasting management regimes. Proceedings of the National Academy of Sciences, 113(18), 5125–5129. https://doi.org/10.1073/pnas.1520420113

Fisheries management indicators
> Melnychuk, M. C., Peterson, E., Elliott, M., & Hilborn, R. (2017). Fisheries management impacts on target species status. Proceedings of the National Academy of Sciences, 114(1), 178–183. https://doi.org/10.1073/pnas.1609915114

IUU Consolidated vessel list
> NEED

Cap and Tier Data from the US Proposal
> Citation for this?

@@

```{r}
# Load country dependencies helper file
country_dependencies <- read_csv(here::here("data", "lookup-tables", "country_dependencies.csv"))

>>>>

```{r}
### Paths to demographic data files
# WDI data - no path, accessed through package
fao_fishers_path <- here::here("data", "raw", "FAO-2019-yearbook", "fao_fishers_yearbook_070619.csv")
teh_fte_path <- here::here("data", "raw", "Teh-and-Sumaila-2013-fte", "Fisher employment FTE.xlsx")

### Paths to fisheries subsidies data files
sumaila_subsidies_path <- here::here("data", "raw", "Sumaila-et-al-2019-subsidies", "Subsidies_PEW_Final.xlsx")
oecd_fse_path <- here::here("data", "raw", "OECD-2019-fse", "FISH_FSE_10032020183349581.csv")
schuhbauer_subsidies_ssf_path <- here::here("data", "raw", "Schuhbauer-et-al-2017-ssf-subsidies", "Subsidies_SSF_split_2018.csv")

### Paths to other fisheries data files
fao_production_path <- here::here("data", "raw", "FAO-2019-capture-production", "Capture_2019.1.0/")
#fao_value_path
melnychuk_price_path <- here::here("data", "raw", "Melnychuk-et-al-2013-price", "Melnychuk_Exvessel_Price_Database.csv")
ram_sadb_path <- here::here("data", "raw", "RAM-2018-stock-assessment-database", "RLSADB v4.40", "DB Files With Assessment Data", "RLSADB v4.40 (assessment data only).xlsx")
costello_upsides_path <- here::here("data", "raw", "Costello-et-al-2016-upsides", "ProjectionData_2012.csv")
melnychuk_fmi_path <- here::here("data", "raw", "Melnychuk-et-al-2017-fmi", "fmi_values_country.csv")
iuu_vessel_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")

### Paths to other necessary files
cap_tier_path <- here::here("data", "raw", "USA-Cap-Tier-Proposal", "US_cap_tier_data.csv")
country_lookup_path <- here::here("data", "lookup-tables", "country_dependencies.csv")
fao_regions_for_ram_lookup_path <- here::here("data", "lookup-tables", "fao_areas_for_ram_stocks_v4.491.csv")
```

@@

@@

Total number of fishers by state from the FAO Yearbook of Fishery and Aquaculture Statistics (misc. years between 1995 and 2017). These data were manually transcribed from PDF (`./data/raw/fleet_&_employment.pdf`) to CSV format on July 6, 2019 and are stored as `r fao_fishers_path`. 
> FAO. (2019). FAO Yearbook of Fishery and Aquaculture Statistics 2017 (p. 108).

Estimated number of full-time jobs in fisheries from Teh and Sumaila (2013) in 2003. These data were provided to us directly by the authors and are stored as `r teh_fte_path`. 
> Teh, L. C. L., & Sumaila, U. R. (2011). Contribution of marine fisheries to worldwide employment. Fish and Fisheries, 14(1), 77–88. https://doi.org/10.1111/j.1467-2979.2011.00450.x

**Fisheries Subsidies Data**
Estimates of annual fisheries subsidies provision by state from Sumaila et al. (2019) in 2018 US\$. These data were provided to us directly by the authors and are stored at `r sumaila_subsidies_path`.
> Sumaila, U. R., Ebrahim, N., Schuhbauer, A., Skerritt, D., Li, Y., Kim, H. S., … Pauly, D. (2019). Updated estimates and analysis of global fisheries subsidies. Marine Policy, 109, 103695. https://doi.org/10.1016/j.marpol.2019.103695

Estimates of fisheries support provision by state from the OECD (2008 - 2018). These data were last downloaded from the OECD iLibrary portal on March 10, 2019 and are stored as `r oecd_fse_path`.
> OECD. (2019). Fisheries: Fisheries support estimates. Fisheries and Aquaculture Statistics, OECD Agriculture Statistics (Database). https://doi.org/10.1787/ade64fdc-en

Estimates of annual fisheries subsidies provision to small scale fisheries by state in 2018 US\$. Data used in this analysis are unpublished updates (based on the Sumaila et al. (2019) data) from those reported in Schuhbauer et al. (2017) (which are in 2009 US\$). These data were provided to us directly by the authors and are stored at `r schuhbauer_subsidies_ssf_path`. 
> Schuhbauer, A., Chuenpagdee, R., Cheung, W. W. L., Greer, K., & Sumaila, U. R. (2017). How subsidies affect the economic viability of small-scale fisheries. Marine Policy, 82, 114–121. https://doi.org/10.1016/j.marpol.2017.05.013

**Other Fisheries Data**
Fisheries landings in 2018 from the FAO Global Capture Production database (2020). These data were last downloaded from the FAO website on July 25, 2020 and are stored as `r fao_production_path`. 
> FAO. (2020). Global Capture Production Dataset. http://www.fao.org/fishery/statistics/global-capture-production/en

Estimates of ex-vessel prices created by Melnychuk et al. (2017). These data were provided to us directly by the authors and are stored as `r melnychuk_price_path`. 
> Melnychuk, M. C., Clavelle, T., Owashi, B., & Strauss, K. (2017). Reconstruction of global ex-vessel prices of fished species. ICES Journal of Marine Science, 74(1), 121–133. https://doi.org/10.1093/icesjms/fsw169

The RAM Legacy Stock Assessment Database is a compolation of stock assessments for commercially exploited species from around the world. The latest version of this database (version 4.491) was downloaded directly from the RAM website on July 25, 2020 and are stored as `r ram_sadb_path`. 
> RAM Legacy Stock Assessment Database. (2020). Version 4.491-assessment only. Released 2020-02-20. https://doi.org/10.5281/zenodo.3676087

Estimates of stock status for comericlaly exploited species not included in the RAM Legacy Stock Assessment Database from Costello et al. (2016). These data were provided to us directly by the authors and are stored as `r costello_upsides_path`. 
> Costello, C., Ovando, D., Clavelle, T., Strauss, C. K., Hilborn, R., Melnychuk, M. C., Branch, T. A., Gaines, S. D., Szuwalski, C. S., Cabral, R. B., Rader, D. N., & Leland, A. (2016). Global fishery prospects under contrasting management regimes. Proceedings of the National Academy of Sciences, 113(18), 5125–5129. https://doi.org/10.1073/pnas.1520420113

Fisheries management indicators for many of the top fishing nations from Melnychuk et al. (2017). These data were manually transcribed from Figure S2 in the supplementary information and are stored as `r melnychuk_fmi_path`. 
> Melnychuk, M. C., Peterson, E., Elliott, M., & Hilborn, R. (2017). Fisheries management impacts on target species status. Proceedings of the National Academy of Sciences, 114(1), 178–183. https://doi.org/10.1073/pnas.1609915114

The IUU Consolidated Vessel List created and maintained by Trygg Mat Tracking. These data were downloaded directly from the website on November 30, 2019 and are stored as `r iuu_vessel_path`. 
> Trygg Mat Tracking. (2020). Combined IUU Fishing Vessel List. https://www.iuu-vessels.org/

We also use the following tables to aid in our analysis:

- Proposed tier assignments for a cap-based approach to limiting subsidies as included in the proposal made to the WTO by the US Delegation on May 3, 2019 [TN/RL/GEN/197/Rev.1].

- A manually compiled table of overseas dependencies of WTO Member and Observer countries

- A manually compiled table of FAO statistical regions coorepsonding to the stocks contained in version 4.491 of the RAM Legacy Stock Assessment Database.

@@

```{r}
# Load country dependencies helper file
country_dependencies <- read_csv(country_lookup_path)
